\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-09-17 overview.tex
update

We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing. We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}). Furthermore, the language features that are required to be used to connect the real code with the modeled code (e.g., \emph{virtual method dispatch}), are already being heavily used in production for testing purposes , which significantly lowers the bar for practitioners to embrace \psharp for testing.

@@

{\bf Safety}.
Safety property specifications can be encoded in \psharp by using the provided \texttt{Assert(...)} method. This method takes as argument a predicate, which if evaluates to false denotes a safety property violation. The programmer is also free to use other custom assertion APIs (as \psharp executes actual code), but to enable \psharp to recognize such APIs, the suggested approach is to override them to call the \psharp \texttt{Assert(...)} method.

\psharp also provides a way to specify global assertions by using \emph{monitors}, special state machines that
can only receive events, but not send. To declare a monitor, the programmer has to inherit from the \psharp \texttt{Monitor} class. A \texttt{Machine} does not need to have a reference to a \texttt{Monitor} to send an event to it; as long as a \texttt{Monitor} has been created, any \texttt{Machine} can invoke it by calling \texttt{Monitor<M>(...)}, where \texttt{M} is the identifier name of the \texttt{Monitor}, and the parameter is an event and an optional payload. This call is also synchronous, in contrast to the regular asynchronous \texttt{Send(...)} method calls, as the monitor has to be called deterministically (and not be interleaved).

{\bf Liveness}.
Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states. Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.

The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states. A state annotated with the \texttt{Hot} attribute denotes a state where the liveness property is not satisfied (e.g. a node has failed but a new one has not come up yet). A state annotated with the \texttt{Cold} attribute denotes a state where the liveness property is satisfied.

One can imagine a liveness monitor as a \emph{thermometer}: as the program executes and the liveness property is not satisfied, the liveness monitor stays in the hot state, which infinitely raises the temperature. If the liveness property is ever satisfied, and thus the liveness monitor transitions to a cold state, the temperature is instantly reduced to normal levels. Encoding liveness properties using hot and cold states enables the programmer to specify arbitrary LTL properties. The liveness monitors in \psharp are based on previous work~\cite{Desai:EECS-2015-198}.

\subsection{Systematic testing}
\label{sec:psharp:testing}
In this section, we present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.

A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.

The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines. A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions. We now give an overview of how this works, while more details can be found in the original paper~\cite{deligiannis2015psharp}.

When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic testing engine} captures and takes control of all sources of non-determinism that are \emph{known} to the \psharp runtime. The engine attempts to detect asynchronous bugs by systematically scheduling machines to execute their event handlers in a different order. This approach is based on systematic concurrency testing~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} (SCT) techniques that have been previously developed for testing shared memory programs.

The systematic testing engine serializes the program execution, takes control of the synchronization and nondeterministic choice points, and at each \emph{step} of the execution it invokes a systematically chosen \psharp machine to execute its next event handler. The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings, until it either reaches a bound (in number of iterations or time), or it hits an assertion failure. The testing is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules.

We have implemented two schedulers inside the \psharp systematic testing engine: \emph{random} and \emph{probabilistic concurrency testing} (PCT)~\cite{burckhardt2010pct}. The random scheduler takes a completely random decision at every scheduling point, whereas the PCT scheduler uses randomization in a disciplined fashion to provide a probabilistic guarantee of finding a bug. It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface exposed by the \psharp libraries. The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers, although we have experimented only with generic schedulers so far. Exposing an easy-to-use interface for creating new schedulers is inspired by previous work~\cite{DesaiQS15}.

We designed the systematic testing engine in a way that enables easy debugging: after a bug is found, the engine can generate a trace that represents the buggy schedule. Note that this trace (in contrast to typical logs generating during production) is sequential.

>>>>

We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing. We also argue that our approach is \emph{general} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:cases}). Furthermore, the language features that are required to be used to connect the real code with the modeled code (e.g., \emph{virtual method dispatch}), are already being heavily used in production for testing purposes , which significantly lowers the bar for practitioners to embrace \psharp for testing.

@@

{\bf Safety}.
In  addition to the usual assertions useful for stating local safety properties of a machine, 
\psharp also provides a way to specify global assertions by using a \emph{safety monitor},
which is a special state machine that can receive but not send events.
A safety monitor maintain local state that is modified in response to messages received from ordinary machines.
This local state is used to maintain a history of the computation that is relevant to the property being specified.
An erroneous global behavior is flagged via an assertion on the private state of the monitor.

To declare a monitor, the programmer inherits from the \psharp \texttt{Monitor} class.
A machine does not need a reference to a \texttt{Monitor} object to send an event to it.
If an instance of \texttt{M}, a class that inherits from \texttt{Monitor}, has been created,
a machine can invoke it by calling \texttt{Monitor<M>(e)},
where the parameter \texttt{e} is an event together with an optional payload.

{\bf Liveness}.
Liveness property specifications are more difficult to encode since they are intended to capture program progress
over infinite executions.
Violation of a liveness property can only be demonstrated by an infinite execution.
Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
We take a different approach and allow the programmer to write a \emph{liveness monitor}.
Similar to a safety monitor, a liveness monitor receives events but does not send them.
Unlike a safety monitor, a liveness monitor~\cite{Desai:EECS-2015-198}
contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
A hot state denotes a point in the execution where progress is required but has not happened yet,
e.g., a node has failed but a new one has not come up yet.
A liveness monitor enters a hot state upon receiving notification of an event that requires the system
to make progress and leaves the hot state upon receiving another event notifying it of progress.
An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
Our liveness monitors can encode arbitrary temporal logic properties.

A thermometer analogy is useful for understanding a liveness monitor.
A hot state increases the temperature by a small value.
A cold state resets the temperature to zero.
A liveness error happens if the temperature becomes infinite.
%The distinction between ordinary and cold states is useful for modeling progress for each instance of an infinite stream of events,
%such as a periodic request or a periodic timer expiration.

\subsection{Systematic testing}
\label{sec:psharp:testing}

The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines. A key capability of the \psharp runtime is that it can execute in bug-finding mode for systematically testing a \psharp program to find bugs, such as assertion violations and uncaught exceptions. We now give an overview of how this works; more details can be found in a previous paper~\cite{deligiannis2015psharp}.

When the \psharp runtime runs in bug-finding mode, an embedded \emph{systematic concurrency testing engine}
~\cite{godefroid1997verisoft, musuvathi2008finding, emmi2011delay} captures and takes control of all sources
of non-determinism that are \emph{known} to the \psharp runtime.
In particular, the runtime is aware of nondeterminism due to interleaving of event handlers in different machines.
Each send operation to an ordinary (non-monitor) machine and each create opetion of an ordinary (non-monitor) machine creates an
interleaving point where the runtime makes a scheduling decision.
The engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of interleavings,
until it either reaches a bound (in number of iterations or time), or it hits a property violation.
The testing is fully automatic, has no false-positives (assuming an accurate environmental model),
and can reproduce found bugs by replaying buggy schedules.
After a bug is found, the engine generates a trace that represents the buggy schedule.
In contrast to typical logs generating during production, this log provides a global order of all communication events
and is easy to debug.

We have implemented two schedulers inside the \psharp systematic testing engine:
a \emph{random} scheduler that makes each nondeterministic choice randomly
and \emph{PCT} scheduler~\cite{burckhardt2010pct}.
It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{DesaiQS15}
exposed by the \psharp libraries.
The interface exposes callbacks that are invoked by the \psharp runtime for taking decisions regarding which machine to schedule next,
and can be used for developing both generic and application-specific schedulers.

Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
